---
title: "Denovo Chimera Detection"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Denovo Chimera Detection}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Overview

The `rchime()` function allows you to detect and remove chimeric sequences using the dataset as it's own reference (denovo). The denovo approach is our preferred method for removing chimeras.

```{r setup}
library(rchime)
```

## Creating *strollur* objects

The `rchime()` function takes a [strollur::dataset](https://mothur.org/strollur/reference/dataset.html) object as input for the data parameter. The *rchime* package contains sequence data from [mothur's](https://mothur.org) [MiSeq_SOP](https://mothur.org/wiki/miseq_sop/) example analysis. Let's create a dataset from containing those sequences.

```{r}
fasta_data <- readRDS(rchime_example("miseq_fasta.rds"))
abundance_data <- readRDS(rchime_example("miseq_abundance.rds"))

data <- strollur::new_dataset("rchime denovo example")

strollur::add(data, table = fasta_data, type = "sequences")
strollur::assign(data, table = abundance_data, type = "sequence_abundance")

data
```

## Removing chimeras

When removing chimeras using the denovo method, the potential parents are chosen from more abundant sequences in your dataset. 

Before we remove the chimeras let's discuss the *dereplicate* parameter. If your dataset includes sample assignments, we recommend setting the *dereplicate* parameter to `TRUE`. When dereplicate = `FALSE`, if a sequence is flagged as chimeric in one sample, it is removed from all samples. Our experience suggests that this is a bit aggressive since we’ve seen rare sequences get flagged as chimeric when they’re the most abundant sequence in another sample. For a more conservative approach, set dereplicate = `TRUE` which will only remove sequences from the samples in which they are flagged as chimeric. Let's use the denovo method to remove the chimeras.

```{r}
results <- rchime(data, dereplicate = TRUE)

data
```


Let's take a look at the first 5 chimeric sequences in the chimera_report:

```{r}
head(results$chimera_report[results$chimera_report$Chimeric_Status != "N", ],
  n = 5
)
```

