---
title: "Getting Started"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The *rchime* package allows you to detect and remove chimeras from your dataset using a de novo approach or alternatively a reference model. This package uses 
code from the [vsearch](https://github.com/torognes/vsearch) tools.

## Installing rchime

You can install the CRAN version with:

```{r, eval = FALSE}
install.packages("rchime")
```

You can install the development version of *rchime* from [GitHub](https://github.com/mothur/rchime) with:

```{r, eval = FALSE}
pak::pak("mothur/rchime")
```


## Tutorials to help you get started

* [De novo based chimera detection](https://mothur.org/rchime/articles/denovo_based_detection.html)
* [Reference based chimera detection](https://mothur.org/rchime/articles/reference_based_detection.html)

