---
title: "Reference Based Chimera Detection"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Reference Based Chimera Detection}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Overview

The `rchime()` function allows you to detect and remove chimeras from your dataset using a referenced based approach. In this tutorial we use SILVA reference files trimmed to our region of interest. *rchime* is designed to be flexible and you can use any reference you choose. 

```{r setup}
library(rchime)
```

## Creating *strollur* objects

The `rchime()` function takes [strollur::dataset](https://mothur.org/strollur/reference/dataset.html) objects as inputs for the data and reference parameters. Let's create a dataset that contains our reference sequences.

```{r}
reference_data <- readRDS(rchime_example("reference.rds"))

reference <- strollur::new_dataset("Silva V4 Region")
strollur::add(reference, table = reference_data, type = "sequences")

reference
```

The *rchime* package contains sequence data from [mothur's](https://mothur.org) [MiSeq_SOP](https://mothur.org/wiki/miseq_sop/) example. Let's create a dataset from containing those sequences.

```{r}
fasta_data <- readRDS(rchime_example("miseq_fasta.rds"))
abundance_data <- readRDS(rchime_example("miseq_abundance.rds"))

data <- strollur::new_dataset("rchime reference example")

strollur::add(data, table = fasta_data, type = "sequences")
strollur::assign(data, table = abundance_data, type = "sequence_abundance")

data
```

## Removing chimeras

When removing chimeras using a reference, the potential parents are chosen from the set of reference sequences. Let's use the reference to remove the chimeras.

```{r}
results <- rchime(data, reference = reference)

data
```

